---
title: 'Analysis  of model building blocks equivalency'
author:
- name: František Kalvas
  url: https://github.com/frantisek901/Spirals/Experiment
  affiliation: Department of Sociology, University of West Bohemia in Pilsen
  affiliation_url: https://les.zcu.cz
date: "2023-08-28"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    code_folding: hide
  word_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
## Encoding: UTF-8
## Created:  2023-02-09 FrK
## Edited:   2023-08-28 FrK

knitr::opts_chunk$set(echo = TRUE)
```

# Edits  

## 2023-08-28

I spotted reversed order in data on Steps 2&3, now is the order correct, so Step 3 sometimes produces higher polarization than Step 2.  
  
  

# Introduction  

Now let us read data in! 


```{r message=FALSE}
# Clearing all
rm(list=ls())

# Packages:
library(tidyverse)
library(readr)
library(dplyr)
library(tibble)
library(forcats)
library(ggplot2)
library(rstatix)
library(stringr)
library(knitr)
# library(kableExtra)
library(stargazer)
library(caTools)
library(ROCR)


#### Reading data in:
# source("SOM_loading.R")
load("SOM.RData")

```



# Step 1 VS. Step 2  

Firstly, data filtering:   

```{r step 1 vs 2}
df = tc %>% filter((Step == "No Identity, Constant Boundary" | Step == "No Identity, Normal Boundary"), 
                   Boundary_STD == 0, Conformity_STD == 0, HK_distribution == "FALSE")
df %>% count(Step)

```

Comparison:   

Computing sim-by-sim difference:

```{r}
# We split data into pairs according all relevant varibles, except STEP:
dp = df %>%  
  pivot_longer(cols = diversity:ESBG) %>% 
  pivot_wider(id_cols = c(HK_distribution:Conformity_STD, name), names_from = Step, values_from = value) %>% 
  rename("S1" = 14, "S2" = 15) %>% 
  mutate(
    Difference = S2 - S1,
    highDifference = abs(Difference) > 0.1) %>% 
  mutate(across(N:name, ~factor(.x)))

# We split data into groups according all relevant varibles, except STEP:
dg = dp %>% group_by(N, Boundary, Conformity, name) %>% 
  summarise(Difference_sd = sd(Difference)) %>% 
  mutate(highDifference = Difference_sd > 0.1)
   
dp %>% 
  ggplot() +
  aes(x = Difference, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1, -0.1)) +
  labs(title = "Absolute pair differences (S2 - S1)") + 
  theme_light()

dg %>% 
  ggplot() +
  aes(x = Difference_sd, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1)) +
  labs(title = "SD of differences (S2 - S1) in groups") +
  theme_light()

mp = lm(Difference ~ name+Conformity+Boundary+N, data = dp)
mpa = lm(abs(Difference) ~ name+Conformity+Boundary+N, data = dp)
mg = lm(Difference_sd ~ name+Conformity+Boundary+N, data = dg)
stargazer(mp, mpa, mg, type = "text")

```


## Tables with extremes  


```{r}
dp %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme absolute differences in pairs comparison")

dg %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme SD of within-groups comparison")

```


 

# Step 2 VS. Step 3  

Firstly, data filtering:   

```{r step 3 vs 2}
df = tc %>% filter((Step == "Constant SPIRO, Normal Boundary" | Step == "No Identity, Normal Boundary"), SPIRO_Mean == 0.25)
df %>% count(Step)

```

Comparison:   

Computing sim-by-sim difference:

```{r}
# We split data into pairs according all relevant varibles, except STEP:
dp = df %>% 
  pivot_longer(cols = diversity:ESBG, values_drop_na = T) %>% 
  pivot_wider(id_cols = c(RS, HK_distribution, N, Boundary, Boundary_STD, Conformity, Conformity_STD, name), 
              names_from = Step, values_from = value) %>% 
  rename("S2" = 9, "S1" = 10) %>% 
  mutate(
    Difference = S2 - S1,
    highDifference = abs(Difference) > 0.1,
    asymetryDifference = Difference > 0.25) %>% 
  mutate(across(N:name, ~factor(.x)))

# We split data into groups according all relevant varibles, except STEP:
dg = dp %>% group_by(HK_distribution, N, Boundary, Boundary_STD, Conformity, Conformity_STD, name) %>% 
  summarise(Difference_sd = sd(Difference)) %>% 
  mutate(highDifference = Difference_sd > 0.1)


dp %>% 
  ggplot() +
  aes(x = Difference, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1, -0.1)) +
  labs(title = "Absolute pair differences (S3 - S2)") + 
  theme_light()

dg %>% 
  ggplot() +
  aes(x = Difference_sd, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1)) +
  labs(title = "SD of differences (S3 - S2) in groups") +
  theme_light()

mp = lm(Difference ~ name+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
mpa = lm(abs(Difference) ~ name+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
mg = lm(Difference_sd ~ name+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dg)
stargazer(mp, mpa, mg, type = "text")

```

## Extra logistic regression  

```{r}
# Logit regression
ml = glm(asymetryDifference ~ Boundary + Boundary_STD + Conformity + Conformity_STD + N + HK_distribution, 
         data = dp, family = "binomial") 
# stargazer(ml, type = "text")

# Testing ability of prediction
## Preparing splitted file 
splited = dp %>% 
  arrange(Boundary, Boundary_STD, Conformity, Conformity_STD, N, HK_distribution, name) %>% 
  mutate(Test = TRUE)

set.seed(20230828)
for (set in 0:4031) {
  tst = sample(c(rep(F, 48), rep(T, 12)), 60, replace = F)
  splited[((set*60) + 1):((set + 1) * 60), "Test"] = tst
}

mdl = glm(asymetryDifference ~ Boundary + Boundary_STD + Conformity + Conformity_STD + N + HK_distribution, 
         data = filter(splited, !Test), family = "binomial") 

## Comparison of models estimated on full/reduced data:
stargazer(ml, mdl, type = "text", title = "Comparison of models estimated on full/reduced data")

## Calculating moddel precision:
prd = predict(mdl, filter(splited, Test), type = "response") > 0.4
splited$Prediction = NA
splited[splited$Test,]$Prediction = prd

## Quality of prediction of non/extreme
mean(splited[splited$asymetryDifference,]$asymetryDifference == splited[splited$asymetryDifference,]$Prediction,
  na.rm = T) %>% round(., 3) %>% paste("Prediction of extreme:", . * 100, "%") %>% print()
mean(splited[!splited$asymetryDifference,]$asymetryDifference == splited[!splited$asymetryDifference,]$Prediction,
  na.rm = T) %>% round(., 3) %>% paste("Prediction of NONextreme:", . * 100, "%") %>% print()
mean(splited[splited$Prediction,]$asymetryDifference == splited[splited$Prediction,]$Prediction,
  na.rm = T) %>% round(., 3) %>% paste("Predicted extreme:", . * 100, "%") %>% print()
mean(splited[!splited$Prediction,]$asymetryDifference == splited[!splited$Prediction,]$Prediction,
  na.rm = T) %>% round(., 3) %>% paste("Predicted NONextreme:", . * 100, "%") %>% print()

```




## Tables with extremes  


```{r}
dp %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(asymetryDifference)) %>% 
  kable(caption = "Table with fraction of extreme positive (asymetric) absolute differences in pairs comparison")

dp %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme absolute differences in pairs comparison")

dg %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme SD of within-groups comparison")

dp %>% filter(RS %in% c(1:2, 6:7, 11, 16, 21, 26)) %>% 
  count(highDifference) %>% kable(caption = "Extreme cases on same seeds: Step 2&3")

```

## Comparison with specially simulated data (SPIRO_Mean==0.001)  

Preparation of data: 

```{r data preparation: Step 2 & specially simulated Step 3, message=F, warning=F}

s2 = tc %>% filter(Step == "No Identity, Normal Boundary") %>% 
  mutate(SPIRO_Mean = 0,
         Step = "Step 2 regular",
         across(c(1, 3:12), ~factor(.x)))  
  

raw3 = read_csv("DATA/Eqivalency2&3_RS01-05-table.csv", skip = 6) %>%
  add_row(read_csv("DATA/Eqivalency2&3_RS06-10-table.csv", skip = 6))
for (i in seq(11, 26, 5)) {
  raw3 = raw3 %>%
    add_row(read_csv(paste0("DATA/Eqivalency2&3_RS", i, "-", i + 4, "-table.csv"), skip = 6))
}

t3 = raw3 %>%
  # Selecting and renaming...
  select(HK_distribution = 4, 2, Use_identity = 12, 13:16,
         N = 3, Boundary = 7, 8, Conformity = 10, 11,
         39, diversity = 40, extremness = 41, ESBG = 42) %>%
  # Dropping duplicate observations due to experiment runs dubling:
  distinct() %>% 
  mutate(Step = "Step 3 check",
         across(c(1, 3:12), ~factor(.x)))

df = s2 %>% add_row(t3)

paste("Simulations from Step 3:", nrow(t3), "what is", round(nrow(t3) / nrow(s2) * 100, 1), "%.") %>% print()
t3 %>% count(RS)

```


Comparison:   

Computing sim-by-sim difference:

```{r}
# We split data into pairs according all relevant varibles, except STEP:
dp = df %>% 
  pivot_longer(cols = diversity:ESBG, values_drop_na = T) %>% 
  pivot_wider(id_cols = c(RS, HK_distribution, N, Boundary, Boundary_STD, Conformity, Conformity_STD, name), 
              names_from = Step, values_from = value) %>% 
  drop_na() %>% 
  rename("S2" = 9, "S3" = 10) %>% 
  mutate(
    Difference = S3 - S2,
    highDifference = abs(Difference) > 0.1,
    asymetryDifference = Difference > 0.25) %>% 
  mutate(across(N:name, ~factor(.x)))

dp %>% 
  ggplot() +
  aes(x = Difference, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1, -0.1)) +
  labs(title = "Absolute pair differences (S3 - S2)") + 
  theme_light()

mp = lm(Difference ~ name+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
mpa = lm(abs(Difference) ~ name+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
stargazer(mp, mpa, type = "text")

dp %>% filter(RS %in% c(1:2, 6:7, 11, 16, 21, 26)) %>% 
  count(highDifference) %>% kable(caption = "Extreme cases on same seeds: Step 2&3")

```

 

# Step 4 VS. Step 3  

Firstly, data filtering:   

```{r step 3 vs 4}
df = tc %>% filter((Step == "Constant SPIRO, Normal Boundary" | Step == "Normal SPIRO, Normal Boundary"), SPIRO_STD == 0)
df %>% count(Step)

```

Comparison:   

Computing sim-by-sim difference:

```{r}
# We split data into pairs according all relevant varibles, except STEP:
dp = df %>% 
  pivot_longer(cols = diversity:ESBG, values_drop_na = T) %>% 
  pivot_wider(id_cols = c(RS, HK_distribution, N, Boundary, Boundary_STD, Conformity, Conformity_STD, SPIRO_Mean, name), 
              names_from = Step, values_from = value) %>% 
  rename("S1" = 10, "S2" = 11) %>% 
  mutate(
    Difference = S2 - S1,
    highDifference = abs(Difference) > 0.1) %>% 
  mutate(across(N:name, ~factor(.x)))

# We split data into groups according all relevant varibles, except STEP:
dg = dp %>% group_by(HK_distribution, N, Boundary, Boundary_STD, Conformity, Conformity_STD, SPIRO_Mean, name) %>% 
  summarise(Difference_sd = sd(Difference)) %>% 
  mutate(highDifference = Difference_sd > 0.1) %>% ungroup()

dp %>% 
  ggplot() +
  aes(x = Difference, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1, -0.1)) +
  labs(title = "Absolute pair differences (S4 - S3)") + 
  theme_light()

dg %>% 
  ggplot() +
  aes(x = Difference_sd, fill = name) +
  facet_wrap(~name) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = c( 0.1)) +
  labs(title = "SD of differences (S4 - S3) in groups") +
  theme_light()

mp = lm(Difference ~ name+SPIRO_Mean+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
mpa = lm(abs(Difference) ~ name+SPIRO_Mean+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dp)
mg = lm(Difference_sd ~ name+SPIRO_Mean+Conformity+Conformity_STD+Boundary+Boundary_STD+N+HK_distribution, data = dg)
stargazer(mp, mpa, mg, type = "text" )


```
  
   
  
## Tables with extremes  

```{r}
dp %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme absolute differences in pairs comparison")

dg %>% group_by(name) %>% summarise(fractionOfHighDifferences = mean(highDifference)) %>% 
  kable(caption = "Table with fraction of extreme SD of within-groups comparison")

```





