---
title: "Report for Ashley"
author: "František Kalvas"
date: "2022-03-07"
output: pdf_document
---

```{r setup, include=FALSE}
## Encoding: UTF-8

rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data

Data are at <http://github.com/frantisek901/Spirals/Experiment>. Experiment is still running and I, Francesco, from time to time actualize the `*.csv` files at GitHub, then I run script `experiment.R` which loads the data. Later version probably finds better names for variables, but now, I use default names from NetLogo experiment.

Who is not interested in working with megabytes of `*.csv files`, might use compiled `*.RData`, there are two files: `shortData.RData`, which is main data file from experiments running only 365 steps, these data are extended by extra simulations with low size of small-world network neighborhood; and `longData.RData`, which is additional data file from experiments running 3650 steps -- thanks to it we might test the effect of simulation length.

Now we load these data:
```{r loading}
load("shortData.RData")
load("longData.RData")

```



## Regressions



```{r regression, echo=FALSE}
# 365 ticks
summary(lm(normalized_polarization_final~boundary+mode+id_threshold+`use_identity?`+`tolerance-level`+`p-speaking-level`+`conformity-level`+`p-random`+`n-neis`,res)) 
summary(lm(ESBSG_polarization_final~boundary+mode+id_threshold+`use_identity?`+`tolerance-level`+`p-speaking-level`+`conformity-level`+`p-random`+`n-neis`,res))

# 3650 ticks
summary(lm(normalized_polarization_final~boundary+mode+id_threshold+`use_identity?`+`tolerance-level`+`p-speaking-level`+`conformity-level`+`p-random`+`n-neis`, long))
summary(lm(ESBSG_polarization_final~boundary+mode+id_threshold+`use_identity?`+`tolerance-level`+`p-speaking-level`+`conformity-level`+`p-random`+`n-neis`, long))


```

