---
title: "Report for Ashley"
author: "František Kalvas"
date: '2022-03-11'
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
## Encoding: UTF-8

rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

## Packages etc.

```{r}
library(stargazer)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lmtest)
library(forcats)


# My own functon for renaming in Tidyverse
prejmenuj = function(data, positions, new.names) {
  names(data)[positions] = new.names
  data
}
```



## Loading data

Data are at <http://github.com/frantisek901/Spirals/Experiment>. Experiment is still running and I, Francesco, from time to time actualize the `*.csv` files at GitHub, then I run script `experiment.R` which loads the data. Later version probably finds better names for variables, but now, I use default names from NetLogo experiment.

Who is not interested in working with megabytes of `*.csv files`, might use compiled `*.RData`, there are two files: `shortData.RData`, which is main data file from experiments running only 365 steps, these data are extended by extra simulations with low size of small-world network neighborhood; and `longData.RData`, which is additional data file from experiments running 3650 steps -- thanks to it we might test the effect of simulation length.

Now we load these data:
```{r loading}
load("shortData.RData")
load("longData.RData")

```



## Regressions

On the two following pages, there are 4 regressions in 2 tables (I'm starting with `stargazer`, later I will produce better output, but for now...). The first table uses ESBG polarization measure, after 365 and 3650 steps, the second uses my normalized polarization measure after same number of steps.

\pagebreak

```{r regression1, echo=FALSE}
# 365 ticks
nn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
ne = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res))

# 3650 ticks
ln = (lm(normalized_3650~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`, long))
le = (lm(ESBG_3650~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`, long))

stargazer(ne, le, type = "text")

```

\pagebreak

```{r regression2, echo=FALSE}
stargazer(nn, ln, type = "text")

```

### Note:

1. Variables `mode:vaguely-speak` and `use_identity?` are binary, `n-neis` is measured on scale 1--64, and all other variables (`id_threshold`, `boundary` etc.) are measured on scale 0--1.

2. I check the problem of `use_identity?` -- I estimated same regression model on sub-sample of simulation with `use_identity?==TRUE`, naturally, effect of mere `use_identity?` is not estimable, but good news is that effect of `id_threshold` is completely same (OK, up to 5th decimal place).

3. Just for curiosity I estimated the model for subsample `use_identity?==FALSE`, I was surprised that all effects were roughly by one order lower ($10^{-1}$).


## Graphs  

### Sampling  

I produced graphs after some random sampling. Both files standard (365 steps) and long (3650 steps) are huge with many thousands of observations. So I created two samples, each of 10,000 observations -- 5,000 simulations using identity, 5,000 not using identity.

```{r sampling}

res_sample = sample_n(res[res$`use_identity?`,], 5000) %>%
  add_row(sample_n(res[!res$`use_identity?`,], 5000)) %>%
  sample_n(10000)

long_sample = sample_n(long[long$`use_identity?`,], 5000) %>%
  add_row(sample_n(long[!long$`use_identity?`,], 5000)) %>%
  sample_n(10000)

```



### Inter-quartile range

Here we look at depiction of distribution of interquartile range of both opinions. The first graph is made from standard (365 steps) data, the second from long (3,650 steps) data.   

```{r graph1, fig.width=5.8}

res_sample %>% 
  ggplot(aes(x = iqr_op1_final, y = iqr_op2_final, col = `use_identity?`)) +
  geom_point(alpha = 0.35) +
  labs(title = "IQR of both opinions: after 365 steps", 
       caption = "Sample of 5,000 simulations using udentity and 5,000 simulations not using identity.") +
  theme_minimal()



```



```{r graph2, fig.width=5.8}

long_sample %>% 
  ggplot(aes(x = iqr_op1_final, y = iqr_op2_final, col = `use_identity?`)) +
  geom_point(alpha = 0.35) +
  labs(title = "IQR of both opinions: after 3,650 steps", 
       caption = "Sample of 5,000 simulations using udentity and 5,000 simulations not using identity.") +
  theme_minimal()

```

For me the basic logic is same in both graphs: some part of simulations ends up with consensus, mainly its simulations not using identity (red dots). Simulation using identity (turquoise dots) sometimes ends up with consensus as well, but also frequently ends up polarized, which is reflected by turquoise 'perimeter'. It seems to me that this basic logic -- identity use = perimeter of discord -- is same regardless the length of simulation.   
  
But different is cleanness of this pattern. In long data (3,650 steps) it is very clear and there are almost no observations between 'red consensus dot' in left down corner and 'turquoise discord perimeter'.  In standard data (365 steps) there are some observations and the perimeter seems fatter. The result is obvious: some standard simulations (365 steps) ended too early, because their 'longer twins' moved from 'discord perimeter' or space in between to 'concensus dot'. So, let's check the differences in polarization between standard and long data:

```{r graph3, fig.width=5.8}
df = res %>% mutate(file = "standard") %>% 
  rename(ESBG = ESBG_365, normalized = normalized_365, identity = `use_identity?`) %>% 
  add_row(long %>% mutate(file = "long") %>% 
            rename(ESBG = ESBG_3650, normalized = normalized_3650, identity = `use_identity?`)) %>% 
  group_by(file, identity) %>% 
  summarise(ESBG = mean(ESBG), normalized = mean(normalized)) %>% 
  pivot_longer(cols = c(ESBG, normalized), names_to = "polarization_measure", values_to = "polarization")

ggplot(df, aes(x = file, y = polarization, fill = identity)) +
  facet_wrap(vars(polarization_measure)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in \nlong (3,650 steps) and standard (365 steps) simulations\nby polarization measure and identity use (TRUE/FALSE)", caption = "Full aggregated sample.") +
  theme_minimal()

```


We see that long (3,650 steps) simulation are tiny slightly less polarized than short (365 steps) ones, i.e. on the average, the polarization in further more than 3,000 steps slightly decreases from initial value. We also see that `normalized` measure shows slightly higher polarization than `ESBG`. So, we might be quite confident that the length of simulation doesn't spoil the results that much -- since there is some tiny differences in aggregate results, it makes sense to do further analyses on individual level, i.e. level of individual simulation, and compute and plot how many times polarization increases from 365th to 3,650th step and how much, but for now we see that after 365 steps we received almost same picture as after 3,650 steps.  

But the main difference is obviously whether we use identity process or not -- regardless the level of identity threshold (but note that we simulate it only for values 0.39, 0.49, 0.59, since it is so important parameter, we now could look at it in more detail). So, let's look now graphically in same way on data, as we did in regression tables:

```{r graph4, fig.width=5.8, fig.height=5.8}
df = res %>% mutate(file = "standard") %>% 
  rename(ESBG = ESBG_365, normalized = normalized_365, identity = `use_identity?`) %>% 
  add_row(long %>% mutate(file = "long") %>% 
            rename(ESBG = ESBG_3650, normalized = normalized_3650, identity = `use_identity?`)) %>% 
  group_by(file, id_threshold, identity, boundary, mode) %>% 
  summarise(ESBG = mean(ESBG), normalized = mean(normalized)) %>% 
  pivot_longer(cols = c(ESBG, normalized), names_to = "polarization_measure", values_to = "polarization")

# `tolerance-level`             -0.024***                      -0.025***          
#                                (0.001)                        (0.001)           
#                                                                                 
# `p-speaking-level`            -0.018***                      -0.011***          
#                                (0.002)                        (0.003)           
#                                                                                 
# `conformity-level`            -0.056***                      0.011***           
                               # (0.003)                        (0.003)        

df %>% filter(file == "standard") %>% 
  ggplot(aes(fill = as.factor(id_threshold), y = polarization, x = as.factor(boundary))) +
  facet_grid(cols = vars(identity, polarization_measure), rows = vars(mode)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in standard (365 steps) simulations\nby polarization measure, identity use (TRUE/FALSE)\nidentity threshold (0.39, 0.49, 0.59) and mode (listen/speak)", caption = "Full aggregated sample.") +
  theme_minimal()  +
  theme(legend.position = "bottom")


```


Again same graph, just for better view only simulations using identity.


```{r graph5, fig.width=5.8, fig.height=5.8}
df %>% filter(file == "standard", identity) %>% 
  ggplot(aes(fill = as.factor(id_threshold), y = polarization, x = as.factor(boundary))) +
  facet_grid(cols = vars(identity, polarization_measure), rows = vars(mode)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in standard (365 steps) simulations\nby polarization measure, identity threshold (0.39, 0.49, 0.59) and mode (listen/speak)", caption = "Full aggregated sample.") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

In previous graph we saw that while some polarisation might happen even without using identity (especially with narrower boundaries), more polarized simulations on average are that using identity. Effect of identity threshold is non-linear: in simulations with 'openly listen' mode the main polarization increase is between 0.39 and 0.49 values, in mode 'vaguely speak' between values 0.49 and 0.59 (but generally, the later mode is less polarized). It is also interesting, that in mode 'vaguely speak' with boundary widening the polarization always decreases, but in mode 'openly listen' this happens only for the lowest identity threshold value (0.39), for other threshold values (0.49, 0.59) the polarization stays same with widening of boundary or even very slightly increases! 

The last result is very surprising -- Hegselmann-Krause model usually finds overall concensus and avoids polarization with wider boundary, it's one of basic results. But when we introduce identity, then this old true changes or is contingent on simulation mode (speaking/listening) and identity threshold. The classical HK findings still hold true, but only for 'vaguely speak' mode and low identity threshold values.

## More regressions

Here we compare full models `le`, `ln`, `ne` and `nn` with single varibles models, with full models omiting these single variables and with variables blocks. Let's start with blocks!  

### Identity block  

```{r regression block 1, echo=FALSE}
idn = lm(normalized_365~id_threshold+`use_identity?`,res) 
ide = (lm(ESBG_365~id_threshold+`use_identity?`,res))

stargazer(ide, idn, type = "text")

anova(ne, ide)
anova(nn, idn)

lrtest(ne, ide)
lrtest(nn, idn)

```


### Hegselmann-Krause block  

```{r regression block 2, echo=FALSE}
hkn = lm(normalized_365~boundary+mode+`conformity-level`,res) 
hke = (lm(ESBG_365~boundary+mode+`conformity-level`,res))

stargazer(hke, hkn, type = "text")

anova(ne, hke)
anova(nn, hkn)

lrtest(ne, hke)
lrtest(nn, hkn)

```


### Network block  

```{r regression block 3, echo=FALSE}
ntn = lm(normalized_365~`tolerance-level`+`p-random`+`n-neis`,res) 
nte = (lm(ESBG_365~`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(nte, ntn, type = "text")

anova(ne, nte)
anova(nn, ntn)

lrtest(ne, nte)
lrtest(nn, ntn)

```



### SoS block  

```{r regression block 4, echo=FALSE}
ssn = lm(normalized_365~`p-speaking-level`,res) 
sse = (lm(ESBG_365~`p-speaking-level`,res))

stargazer(sse, ssn, type = "text")

anova(ne, sse)
anova(nn, ssn)

lrtest(ne, sse)
lrtest(nn, ssn)

```

OK, we see, that we might omit SoS and Network blocks, we may focus only on Identity and HK blocks.

## Without identity

```{r regression block 5, echo=FALSE}
# 365 ticks
nidn = lm(normalized_365~boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
nide = (lm(ESBG_365~boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(nide, nidn, type = "text")


anova(ne, nide)
anova(nn, nidn)

lrtest(ne, nide)
lrtest(nn, nidn)


```



## Without HK block

```{r regression block 6, echo=FALSE}
# 365 ticks
nhkn = lm(normalized_365~id_threshold+`use_identity?`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
nhke = (lm(ESBG_365~id_threshold+`use_identity?`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(nhke, nhkn, type = "text")


anova(ne, nhke)
anova(nn, nhkn)

lrtest(ne, nhke)
lrtest(nn, nhkn)


```


```{r blocks 1, echo=FALSE}
hkidn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`,res) 
hkide = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`,res))

stargazer(hkide, hkidn, type = "text")

anova(hkide, hke)
anova(hkidn, hkn)

lrtest(hkide, hke)
lrtest(hkidn, hkn)

anova(hkide, ne)
anova(hkidn, nn)

lrtest(hkide, ne)
lrtest(hkidn, nn)

```

OK, then! Evidently the weaker block (Identity) improves model with stronger block (Hegselmann-Krause) a lot. Other blocks improve model also (statistically), but it is weak improvement. It is evident that present data these blocks (Identity, HK) explain satisfactorily. But just let's look at Network block -- would it improve the present model (Identity + HK)?


```{r blocks 2, echo=FALSE}
hkidnwn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
hkidnwe = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(hkidnwe, hkidnwn, type = "text")

anova(hkide, hkidnwe)
anova(hkidn, hkidnwn)

lrtest(hkide, hkidnwe)
lrtest(hkidn, hkidnwn)

anova(hkidnwe, ne)
anova(hkidnwn, nn)

lrtest(hkidnwe, ne)
lrtest(hkidnwn, nn)

```

OK, Network and SoS improve model, but it is very weak. Thanks to huge number of obsefvations we receive statistically significant results, even in case of SoS, but substantively it's not big improvement (instead of thousands, only tens of point of $\chi^2 $). Only just for sure, we continue with testing effect of individual variables.


## Individual variables

This testing is not very useful, I think, because we just tested effect of SoS -- block of one variable -- and even this weak variable is statistically significant. But, OK, for sure, we do it! :-) And we will start from the tail, from weakest variables. We will estimate model with just one variable, and also just without this one variable.

### Number of neighbors  

```{r regression idividual1, echo=FALSE}
nnein = lm(normalized_365~`n-neis`,res) 
nneie = (lm(ESBG_365~`n-neis`,res))
nnnein = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`,res) 
nnneie = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`,res))

stargazer(nneie, nnein, type = "text")

anova(nnnein, nn)
anova(nnneie, ne)

lrtest(nnnein, nn)
lrtest(nnneie, ne)

```

OK, Size of neighborhood explains data better, improves model significantly, but still it is weak improvement.


### Probability of random links  

```{r regression idividual2, echo=FALSE}
nvarn = lm(normalized_365~`p-random`,res) 
nvare = (lm(ESBG_365~`p-random`,res))
nnvarn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`tolerance-level`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

Whoa! Probability of random links doesn't improve model significantly! This probability is useless.



### Tolerance to opponents in neighborhood  

```{r regression idividual3, echo=FALSE}
nvarn = lm(normalized_365~`tolerance-level`,res) 
nvare = (lm(ESBG_365~`tolerance-level`,res))
nnvarn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

OK, Tolerance to opponents in neighborhood explains data better, improves model significantly, but still it is weak improvement.



### Probability of speaking

```{r regression idividual4, echo=FALSE}
ssn = lm(normalized_365~`p-speaking-level`,res) 
sse = (lm(ESBG_365~`p-speaking-level`,res))
hkidnwn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
hkidnwe = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(sse, ssn, type = "text")

anova(ne, hkidnwe)
anova(nn, hkidnwn)

lrtest(ne, hkidnwe)
lrtest(nn, hkidnwn)

```

OK, Probability of speaking explains data better, improves model significantly, but still it is very weak improvement.



### Conformity level  

```{r regression idividual5, echo=FALSE}
nvarn = lm(normalized_365~`conformity-level`,res) 
nvare = (lm(ESBG_365~`conformity-level`,res))
nnvarn = lm(normalized_365~id_threshold+`use_identity?`+boundary+mode+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+mode+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

OK, Conformity level in neighborhood explains data better, improves model significantly, but still it is weak improvement.



### Mode  

```{r regression idividual6, echo=FALSE}
nvarn = lm(normalized_365~mode,res) 
nvare = (lm(ESBG_365~mode,res))
nnvarn = lm(normalized_365~id_threshold+`use_identity?`+boundary+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+`use_identity?`+boundary+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

WHOOOOAAA! Mode is hugely significant, explains data so much better, improves model very significantly, it is strong improvement.



### Boundary  

```{r regression idividual7, echo=FALSE}
nvarn = lm(normalized_365~boundary,res) 
nvare = (lm(ESBG_365~boundary,res))
nnvarn = lm(normalized_365~id_threshold+`use_identity?`+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+`use_identity?`+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

OK, Boundary explains data better, improves model significantly, but still it is weak improvement.



### Identity use  

```{r regression idividual8, echo=FALSE}
nvarn = lm(normalized_365~`use_identity?`,res) 
nvare = (lm(ESBG_365~`use_identity?`,res))
nnvarn = lm(normalized_365~id_threshold+boundary+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~id_threshold+boundary+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

WHOOOOAAA! Identity use is very significant, explains data much better, improves model significantly, it is strong improvement.



### Identity level  

```{r regression idividual9, echo=FALSE}
nvarn = lm(normalized_365~id_threshold,res) 
nvare = (lm(ESBG_365~id_threshold,res))
nnvarn = lm(normalized_365~`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res) 
nnvare = (lm(ESBG_365~`use_identity?`+boundary+mode+`conformity-level`+`tolerance-level`+`p-speaking-level`+`p-random`+`n-neis`,res))

stargazer(nvare, nvarn, type = "text")

anova(nnvarn, nn)
anova(nnvare, ne)

lrtest(nnvarn, nn)
lrtest(nnvare, ne)

```

WHOOOOAAA! Identity level is very significant, explains data much better, improves model significantly, it is strong improvement.  

OK, then! So, it seems that there are only 3 variables really significant, all explaining 10+% of variability, they are: Identity use, Identity level and Mode. So for the final regression test, let's compare model with just these three vars and full model, let's see how different these models are.



### Best model?  

```{r blocks 3, echo=FALSE}
idmn = lm(normalized_365~id_threshold+`use_identity?`+mode, res) 
idme = (lm(ESBG_365~id_threshold+`use_identity?`+mode, res))
widmn = lm(normalized_365~boundary+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res) 
widme = (lm(ESBG_365~boundary+`conformity-level`+`p-speaking-level`+`tolerance-level`+`p-random`+`n-neis`,res))

stargazer(idme, ne, type = "text")

stargazer(idmn, nn, type = "text")

stargazer(widme, widmn, type = "text")

anova(idme, ne)
anova(idmn, nn)

lrtest(idme, ne)
lrtest(idmn, nn)

```

OK, resting 6 another variables improves model slightly, explain 1.5 percent points of $R^2$ more. So, yes, the triumvirate variables, 3 the most important, explain almost all variability, the resp explains something, but it's so weak. ANOVA and Likelihood ratio tests tell us that these 6 variables improve, but 67 or 100 points of RSS only. 

So, we might neglect the 6 resting variables in further experiments and play more smoothly with identity threshold (since the Mode is true/false, as well as Identity use). We also should inspect extreme value of Probability of speaking -- if the probability will be 100%, would the results be the same?


## Graph reflecting the best model


```{r graph5, fig.width=5.8, fig.height=5.8}
df = res %>% mutate(file = "standard") %>% 
  rename(ESBG = ESBG_365, normalized = normalized_365, identity = `use_identity?`) %>% 
  add_row(long %>% mutate(file = "long") %>% 
            rename(ESBG = ESBG_3650, normalized = normalized_3650, identity = `use_identity?`)) %>% 
  group_by(file, id_threshold, identity, mode) %>% 
  summarise(ESBG = mean(ESBG), normalized = mean(normalized)) %>% 
  pivot_longer(cols = c(ESBG, normalized), names_to = "polarization_measure", values_to = "polarization")

df %>% #filter(file == "standard") %>% 
  ggplot(aes(x = as.factor(id_threshold), y = polarization, fill = fct_rev(file))) +
  facet_grid(cols = vars(fct_rev(mode), polarization_measure), rows = vars(identity)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in standard (365 steps) simulations\nby polarization measure, identity use (TRUE/FALSE)\nidentity threshold (0.39, 0.49, 0.59) and mode (listen/speak)", caption = "Full aggregated sample.") +
  theme_minimal()  +
  theme(legend.position = "bottom")

df %>% filter(polarization_measure == "ESBG") %>% 
  ggplot(aes(x = as.factor(id_threshold), y = polarization, fill = fct_rev(file))) +
  facet_grid(cols = vars(fct_rev(mode), polarization_measure), rows = vars(identity)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in standard (365 steps) simulations\nby polarization measure, identity use (TRUE/FALSE)\nidentity threshold (0.39, 0.49, 0.59) and mode (listen/speak)", caption = "Full aggregated sample.") +
  theme_minimal()  +
  theme(legend.position = "bottom")

df %>% filter(polarization_measure == "ESBG", file == "standard") %>% 
  ggplot(aes(fill = as.factor(id_threshold), y = polarization, x = fct_rev(mode))) +
  facet_grid(cols = vars(identity)) +
  geom_col(position = position_dodge()) +
  labs(title = "Comparison of average polarization in standard (365 steps) simulations\nby polarization measure, identity use (TRUE/FALSE)\nidentity threshold (0.39, 0.49, 0.59) and mode (listen/speak)", caption = "Full aggregated sample.") +
  theme_minimal()  +
  theme(legend.position = "bottom")



```

